### 2017-08-02 ms.tex
01 Aug mods

%-------------------------------------------------------------------------------
% Kyle Westfall
% westfall@ucolick.org
% UCO/Lick Observatory
% University of California, Santa Cruz
% 1156 High St.
% Santa Cruz, CA 95064
% USA
%
% VERSION:
%       05 Feb 2017: (KBW) v01: stab in the dark
%       01 Aug 2017: (KBW) edits and summary of todo tasks
%
% SUBMITTED:
%
%-------------------------------------------------------------------------------

%\input{../dm_nomenclature/dmnom_rv4.tex}

\documentclass[apj,iop,revtex4,numberedappendix]{emulateapj}
\RequirePackage{calc}
\usepackage{natbib}
\usepackage{url}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{color}
\bibliographystyle{apj}

\slugcomment{Draft: 14 Feb 2017}

% Some fancy commenting
\definecolor{todo}{RGB}{200,0,0}
\newcommand{\comment}[2][todo]{{\color{#1}[[{\bf #2}]]}}

\newcommand{\kms}{{km$~\!$s$^{-1}$}}
\newcommand{\nsaz}{$z_{\rm NSA}$}
\newcommand{\halpha}{H$\alpha$}

\shortauthors{Westfall et al.}
\shorttitle{Asymmetric Drift across the Blue Cloud}

\begin{document}

\title{ The trend in asymmetric drift across the blue cloud }

\author{ Kyle B. Westfall\altaffilmark{1}, Matthew A.
Bershady\altaffilmark{2}, Kevin Bundy\altaffilmark{1}, et al. }

\altaffiltext{1}{UCO/Lick Observatory, University of California, Santa
Cruz, 1156 High St., Santa Cruz, CA 95064, USA}

\altaffiltext{2}{Department of Astronomy, University of Wisconsin-Madison, 475
N. Charter St., Madison, WI 53706, USA}

\email{westfall@ucolick.org}

\begin{abstract}

Asymmetric drift (AD) is the lag of the mean rotation velocity of the
stellar disk behind the circular speed defined by the total
gravitational potential.  Although it is often considered a nuisance
correction one must apply in circular-speed calculations, the direct
connection between AD and the stellar phase-space distribution function
makes it an interesting dynamical properties of galaxies in and of
itself.  The SDSS-IV/MaNGA survey provides more than an order of
magnitude increase in any galaxy sample size useful for AD measurements.
In this pilot Letter, we study the empirical correlation between AD, or
more precisely the differential tangential lag between the stellar
component and \halpha-emitting gas, and global photometric properties
from the NASA-Sloan Atlas.  We isolate a set of galaxies with \halpha\
and stellar velocity fields that are well-fit by simple disk models.
Our sample of $\sim$800 galaxies shows a clear correlation between
absolute $i$-band magnitude and AD measured at half an effective radius.
Removing the primary dependence on the rotation speed, we find that the
stellar disk rotation is 90\% of the \halpha\ rotation speed, in the
mean.  Our sample size allows us to infer a weak, yet statistically
significant, trend with galaxy $N-r$ color such that, in the mean,
redder galaxies have moderately larger AD relative to their rotation
speeds.  Within the context of an albeit strong set of assumptions,
including a direct proportionality between AD and stellar velocity
dispersion as seen in the Milky Way, we argue that our results suggest
little variation in the mean disk mass-to-light ratio as a function of
absolute magnitude and only modest variations with galaxy color
\comment{last sentence TBD; is Figure 3 consistent with SPS
variations?}.

\end{abstract}

\keywords{ galaxies: kinematics and dynamics --- galaxies: spiral ---
galaxies: structure }

\section{ Motivation }
\label{sec:intro}

For ensembles of stars in a galaxy disk, \citet[][Section
4.4.3]{2008gady.book.....B} provide an intuitive description of
asymmetric drift (AD) as arising due to the combined effect of the
radially decreasing surface-density and velocity-dispersion profiles
typical of axisymmetric systems.  Using the $v_R$ moment of the
collisionless Boltzmann equation, one obtains the Jeans \comment{ref}
equation that directly relates the circular speed ($v_c$), the mean
stellar tangential speed ($\overline{v_\theta}$), and the stellar
velocity ellipsoid (SVE) as a function of radius in the plane of
symmetry:
%
\begin{equation}
%
v_c^2 - \overline{v_\theta}^2 = \sigma_R^2\left[
\frac{\sigma_\theta^2}{\sigma_R^2} -
\frac{R}{\rho\sigma_R^2}\frac{\partial(\rho\sigma_R^2)}{\partial R} -
1\right] - R\frac{\partial\overline{v_R v_z}}{\partial z},
%
\label{eq:adformal}
%
\end{equation}
%
where $R,\theta,z$ are the cylindrical coordinates, $\rho$ is the volume
density, and $\sigma$ is the velocity dispersion.  Along with the
standard assumptions of dynamical equilibrium and negligible radial and
vertical flows inherent to its derivation, equation \ref{eq:adformal} is
often simplified by assuming the right-most term --- describing the
covariance between the radial and vertical motions as a function of
perpendicular distance to the plane of symmetry --- is negligible
\comment{cf.  Cuddeford \& Amendt}.

Asymmetric drift has been measured in numerous systems.  Perhaps most
notably, e.g., \comment{Dehnen \& Binney} have shown that populations of
stars show a direct correlation between their velocity dispersion and
the degree to which their mean rotation speed lags behind that of the
Local Standard of Rest (LSR), much in line with the expectation provided
by equation \ref{eq:adformal}.  Although there have been some studies of
AD in external galaxies \comment{refs}, it is often treated as a
nuisance phenomenon that must be corrected for when trying to constrain
the circular speed curve of a galaxy \comment{refs}.  The ubiquity and
phenomenology of AD as a salient observable has not yet been studied for
a statistically significant population of galaxies.  Our aim is
therefore to provide a first look at the correlation between AD
measurements and basic broad-band photometric properties for a large
sample of galaxies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%
\begin{center}
%
\includegraphics[width=0.9\textwidth]{figs/cmd_flux.pdf}
%
\end{center}
%
\caption{
%
${\rm N} - r$ color and absolute $i$-band magnitude ($M_i$) for all
galaxies observed during the first two years of the MaNGA Survey (left),
the subsample of galaxies {\em not} selected to be kinematically regular
(middle; see text), and the AD sample used throughout the remainder of
our analysis.  The color of the data points in the top row represent the
mean \halpha\ surface brightness \comment{check aperture} according to
the colorbar to the right; the bottom row replaces the color by the
$r$-band surface brightness density.
%
}
%
\label{fig:sample}
%
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In addition to the basic understanding of this dynamically relevant
quantity, our interest in AD stems from its fundamental dynamical
connection to the full phase-space distribution function of a galaxy's
stars, as empirically demonstrated in the Milky Way.  Indeed,
\comment{Westfall et al. (2011; see also refs} used AD to constrain the
axial ratios of the SVE given direct measurements of the line-of-sight
(LOS) stellar velocity dispersion, $\sigma$.  Alternatively, if one can
statistically constrain the shape of the SVE \comment{refs},
measurements of AD could be used as a proxy for stellar $\sigma$ by
either appealing to equation \ref{eq:adformal} or an empirical relation
calibrated by samples with both AD and $\sigma$ measurements \comment{in
prep...}.  Such a use of AD as a proxy for stellar $\sigma$ is
attractive in the low-surface-brightness and low-velocity-dispersion
regimes where direct measurements are difficult and/or expensive.  We
commit to this paradigm by using the following nomenclature for AD,
following from equation \ref{eq:adformal}:
%
\begin{equation}
%
\sigma_a^2 \equiv v_c^2 - \overline{v_\theta}^2.
%
\label{eq:addef}
%
\end{equation}

For disk galaxies, there is a decades-long industry of using cold-gas
tracers to construct circular-speed curves of galaxies \comment{refs}.
These data have provided the first concrete arguments for the presence
of massive dark-matter halos \comment{ref} based on mass-model
reconstruction \comment{refs} and have formed the basis for the most
robust Tully-Fisher relations compiled for the local Universe
\comment{refs}.  However, it is important to acknowledge that one cannot
directly measure $v_c$, effectively making $\sigma_a$ unobservable.
That is, every dynamical tracer has some non-zero dynamical pressure
such that it will lag behind the theoretically defined circular speed.
There are very clear examples of early-type galaxies that show
differential AD in their molecular (e.g., CO), atomic (e.g.,
\ion{H}{1}), and/or ionized (\halpha) tracers relative to a robust mass
model of $v_c$ \comment{Davis et al.}, due to the different
turbulent/thermal pressures intrinsic to these tracers.  However, these
signatures are much less apparent in disk galaxies.  For example,
\comment{Martinsson et al.} show that \halpha\ and \ion{H}{1} rotation
curves are consistent for the DiskMass Survey within the limits of their
constraints on beam-smearing, suggesting that any correction for the lag
of either behind the circular speed should be small.  Indeed,
theoretical calculations \comment{Dalcanton \& Stilp} suggest that
gas-pressure corrections to \ion{H}{1} rotation curves should be largely
negligible for galaxies with \comment{${\mathcal M} > 10^x {\mathcal
M}_\odot$}.  Nevertheless, we emphasize here that our measurements of
$\sigma_a$ may be better termed as a {\em differential tangential lag}
because they are based on the quadrature difference between the observed
\halpha\ and stellar rotation curves in our galaxy sample (Section
\ref{sec:data}).  These measurements are perfectly valid in their own
right; however, it is their interpretation in the context of the
theoretical definition of AD that should be questioned in some regimes,
as discussed in Section \ref{sec:discussion}.

We present the data used for our analysis in the following section.  In
Section \ref{sec:results}, we demonstrate that there is a strong
correlation between the absolute $i$-band magnitude of a galaxies and
the strength of its AD measured at half of an effective radius (0.5
$R_{\rm eff}$).  We also illustrate the weak color dependence in this
relation.  We summarize and discuss these results in Section
\ref{sec:discussion}. \comment{flesh out}

\section{Data}
\label{sec:data}

We use integral-field spectroscopy from the SDSS-IV/MaNGA (Mapping
Nearby Galaxies from APO) Survey to construct stellar and ionized-gas
velocity fields for 2715 unique galaxies, 39 of which have multiple
observations.  The kinematic measurements are provided by the MaNGA data
analysis pipeline \comment{DAP; Westfall et al., in prep}: the
ionized-gas kinematics are based on simple single-Gaussian fits to the
\halpha\ emission feature and the stellar kinematics are determined
using {\tt pPXF} (Cappellari et al.) with the MILES stellar template
library \comment{Jesus FB et al.}.

The MaNGA galaxy selection \comment{Wake et al} is a simple cut in
absolute $i$-band magnitude, $M_i$, and redshift, $z$, and provides a
complete \comment{check} sampling of the overall galaxy population.  As
such, measurements of stellar and/or \halpha\ rotation curves for some
systems will be difficult/impossible.  Using the photometry from the
NASA-Sloan Atlas \comment{ref},\footnote{\url{nsatlas.org}} Figure
\ref{fig:sample} shows the color-magnitude diagram of all galaxies
colored by their mean \halpha\ and $r$-band surface brightness.
\comment{ Check the details of how this plot was constructed}  Although
not excluded from our velocity-field analysis {\em a priori}, we expect
to obtain poor rotation-curve measurements for galaxies with low
\halpha\ and stellar surface brightness.\footnote{
%
The uniformity of the MaNGA survey \comment{Law et al. 2015} is such
that the signal-to-noise (S/N) of the datacube spaxels is tightly
correlated with surface brightness.}

\comment{Extinction corrections for the CMD?}

We model the geometric projection of the rotational plane of each galaxy
using the approach presented by \comment{Andersen \& Bershady (2013)};
see also \comment{Westfall et al. (2011)}.  In three independent fitting
iterations, the model fits are optimized for the \halpha\ velocity
field, the stellar velocity field, and simultaneously for both data
sets; for the latter, the geometry is forced to be the same for the two
dynamical tracers, but the parametrized rotation curves are independent.
We use these velocity-field-fitting results to objectively isolate a set
of ``kinematically regular'' galaxies.  Briefly, galaxies in this sample
must have: (i) successful velocity-field fits for all three approaches,
(ii) differences in the measured \halpha\ and stellar systemic velocity
of less than 20 km/s, (iii) dynamical centers that are consistent to
within a fiber diameter of the morphological center \comment{check the
details of this}, (iv) \halpha\ and stellar velocity-field position
angles that are consistent to within $\pm$15$\arcdeg$, and (v) kinematic
inclinations between $15\arcdeg < i < 80\arcdeg$ that are both
consistent between the \halpha\ and stellar data and with respect to the
photometric ellipticity to within $\pm$20$\arcdeg$.  The constraint on
the inclination is by far the most stringent. \comment{give number of
galaxies cut by each criterion?}  Applying these constraints yields a
sample of 798 observations (for 790 unique galaxies) out of the 2764
observations analyzed.  Eleven galaxies with repeat observations satisfy
the selection criteria; however, only five of these show all
observations are consistently selected, whereas the other six show one
or more of the observations did not pass our constraints \comment{we
should understand why repeat observations are not consistent for the
majority of cases.  low S/N?}.  \comment{Do this or remove the
sentence:} Finally, we also visually inspected the broad-band imaging of
these 790 galaxies and eliminated merging and highly extincted (highly
inclined) systems yielding a final sample of XXX galaxies.  We hereafter
refer to galaxies that satisfy our selection criteria as the ``AD
sample''.

Figure \ref{fig:sample} shows the color-magnitude distribution for all
MaNGA galaxies, as well as the distributions of those galaxies included
and excluded from our AD sample.  As expected, galaxies with relatively
high \halpha\ and $r$-band surface brightness are preferentially
selected.  This results in an exclusion of much of the red sequence, as
well as the brightest and faintest galaxies in the blue cloud.  Although
the conclusions we reach based on our AD sample are astrophysically
meaningful, it is important to appreciate that the galaxies in our AD
sample are a biased representation of the overall galaxy population.

\comment{Show histogram of kinematically regular sample against the
volume-corrected distribution of the MaNGA parent sample in Mi and N-r,
and discuss.}

Our velocity-fitting method provides a model rotation curve fit to both
the \halpha\ and stellar data, each parametrized as a hyperbolic tangent
function: $v_{\rm rot} = v_{\rm flat} \tanh(R/h_v)$.  However, our
primary result is based on the error-weighted mean of the deprojected
rotation-curve measurements for each dynamical tracer.  We only include
measurements within $\pm$30$\arcdeg$ of the major axis and construct
radial bins that are $2\farcs5$ wide and centered at 0.25, 0.5, 0.75,
1.0, and 1.25 $R_{\rm eff}$ --- $R_{\rm eff}$ is the effective radius
using the elliptical Petrosian analysis from the NASA-Sloan Atlas.
\comment{Check if the $R_{\rm eff}$ does or does not include the
multiplicative offset to match these Petrosian and Sersic $R_{\rm eff}$
in the mean.}  Specifically, we calculate the error-weighted mean and
standard deviation of $v_j = V_j/\cos\theta_j/\sin i$ and
%
\begin{equation}
%
\sigma_{a,j}^2 = (V_{{\rm H}\alpha,j}^2 - V_{\ast,j}^2) (\cos\theta_j
\sin i)^{-2},
%
\label{eq:dproj}
%
\end{equation}
%
where $V_j$ is the LOS measurement of each component in spaxel $j$
located at the in-plane polar coordinates $R_j$ and $\theta_j$ and $i$
is the disk inclination.

\comment{An example demonstrating the details of our measurements is
illustrated in Figure X.}

We calculate errors in $v_{{\rm H}\alpha}$ and $\sigma_a$ as the
quadrature sum of the error-weighted standard error (i.e., the
error-weighted standard deviation divided by $\sqrt{N}$) and the
propagated error in the error-weighted mean. \comment{Should we revisit
this?  Details of error calculation not all that important.  We're
dominated by intrinsic deviations in the regressions below.}

\section{Results}
\label{sec:results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%
\begin{center}
%
\includegraphics[width=1.0\columnwidth]{figs/mi_models.pdf}
%
\end{center}
%
\caption{
%
Global photometry and kinematic measurements at $R=R_{\rm eff}/2$ for
the AD sample: $M_i$ versus (a) $N-r$ color with each point colored
according to $\sigma_a^2$, (b) $v_{{\rm H}\alpha}$ with each point
colored according to $\sigma_a^2$, (c) $\sigma_a^2$ with each point
colored according to $v_{{\rm H}\alpha}$, and (d) $\sigma_a/v_{{\rm
H}\alpha}$ with each point colored by the $N-r$ color.  Panels (b), (c),
and (d) include the Spearman rank correlation coefficient, $r_s$, and
the linear regressions (solid lines) constructed from the parameters
provided in Table \ref{tab:lines}.  The dashed lines are offset from
linear regression by the modeled intrinsic scatter in the relation
($\varepsilon_y$).
%
}
%
\label{fig:correlation}
%
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Figure \ref{fig:correlation}(a) shows the $(M_i, N-r)$ color-magnitude
diagram (see right column of Figure \ref{fig:sample}) with the points
colored according to $\sigma_a$ measured at $R=0.5R_{\rm eff}$.  There
is a clear trend where brighter galaxies have larger $\sigma_a$; the
remainder of the Figure examines this trend.

We plot $M_i$ versus $v_{{\rm H}\alpha}$, $\sigma_a^2$, and
$\sigma_a/v_{{\rm H}\alpha}$ in Figures \ref{fig:correlation}(b),
\ref{fig:correlation}(c), and \ref{fig:correlation}(d), respectively.
The points are colored according to the label to the right of each color
bar.  Each panel provides the Spearman rank-correlation coefficient,
$r_s$, of the plotted data with errors derived from $10^3$ bootstrap
simulations.  We have also used a Markov Chain Monte Carlo
\comment{refs} to sample the Bayesian posterior distribution for a
linear regression to the data in each panel, incorporating the errors in
both axes \comment{refs}; the errors in the kinematic quantities always
dominate over the absolute magnitude errors.  The fitted model is a line
in parametric form with an intrinsic Gaussian scatter perpendicular to
the line.  That is, the line is defined as $\mathbf{l}(t) = \mathbf{l}_0
+ t\ \hat{\mathbf{l}}$ for a generalized coordinate $t$ along the line,
an origin $\mathbf{l}_0 = \{x_0, y_0\}$, and the unit vector
$\hat{\mathbf{l}} = \{\cos\phi, \sin\phi\}$.  The fitted parameters are
$y_0$, $\phi$, and the dispersion of the intrinsic Gaussian scatter
about the line, $\varepsilon$; $x_0$ is fixed at the median abscissa of
the data being fitted ($x_0 = M_{i,0} = -20.6$).  Uniform priors are
used for $y_0$ and $\phi$ and a logarithmic prior \comment{check this is
true} is used for $\varepsilon$ \comment{ref}.  Using the returned
samples of the posterior, we also provide parameters for the
slope-intercept form of the line --- $y = mx + b$ where $m = \tan\phi$
and $b = y_0 - x_0 \tan\phi$ --- and the scatter projected along the
ordinate, $\varepsilon_y = \varepsilon/|\cos\phi$|.  Table
\ref{tab:lines} provides the median and standard deviation of the
marginalized distribution of each parameter; these parameters have been
used to construct the lines provided in Figure \ref{fig:correlation}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable}{ c r r r }
\tabletypesize{\small}
\tablewidth{0pt}
\tablecaption{Linear Regressions for $M_i$ \comment{check m-b form}}
\tablehead{ & \multicolumn{3}{c}{Dependent Variable} \\ \cline{2-4} \\[-3pt]
 \colhead{Parameter} & \colhead{$\log(v_{{\rm H}\alpha})$} &
 \colhead{$\log(\sigma_a^2)$} &
 \colhead{$\log(\sigma_a/v_{{\rm H}\alpha})$} }
\startdata
          $y_0$ &      2.168 &       3.78 &     -0.269 \\
                & $\pm$0.003 &  $\pm$0.01 & $\pm$0.003 \\[2pt]
         $\phi$ &     170.70 &      160.1 &      -0.24 \\
                &  $\pm$0.15 &   $\pm$0.5 &  $\pm$0.17 \\[2pt]
  $\varepsilon$ &      0.069 &      0.209 &      0.068 \\
                & $\pm$0.001 & $\pm$0.005 & $\pm$0.002 \\[2pt] \hline \\[-4pt]
            $m$ &     -0.164 &     -0.362 &     -0.004 \\
                & $\pm$0.003 & $\pm$0.009 & $\pm$0.003 \\[2pt]
            $b$ &      -1.20 &      -3.69 &      -0.36 \\
                &  $\pm$0.05 &  $\pm$0.19 &  $\pm$0.06 \\[2pt]
$\varepsilon_y$ &      0.070 &      0.222 &      0.068 \\
                & $\pm$0.001 & $\pm$0.005 & $\pm$0.002
\enddata
\label{tab:lines}
\end{deluxetable}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As expected, there is a strong correlation between $v_{{\rm H}\alpha}$
and $M_i$.  However, it is important to note that Figure
\ref{fig:correlation}(b) does not present the Tully-Fisher \comment{ref}
relation for our AD sample; the Figure gives the rotation speed at
$R=0.5R_{\rm eff}$, not a measure of the full-width of the dynamically
broadened line profile.  The slope of the relation in Figure
\ref{fig:correlation}(b) is steeper than a nominal Tully-Fisher relation
because galaxies at low luminosity tend to have more slowly rising
rotation curves \comment{refs}, which can be confirmed by plotting
$h_{v,{\rm H}\alpha}/R_{\rm eff}$ as a function of $M_i$.
\comment{actually show this?}

Figure \ref{fig:correlation}(c) gives the direct representation of the
gradient in the point color seen in Figure \ref{fig:correlation}(a).
The data in this panel are highly correlated, both as determined by
$r_s$ and the fitted regression.  The intrinsic scatter increases from
0.07 dex (17\%) in $(M_i,v_{{\rm H}\alpha})$ to 0.21 dex (62\%) in
$(M_i,\sigma_a^2)$; however, this is close to the $\sim$0.1 dex scatter
if considering $(M_i,\sigma_a)$ instead.

If $\sigma_a$ or $v_{{\rm H}\alpha}$ was a secondary parameter in the
distribution of $(M_i,v_{{\rm H}\alpha})$ or $(M_i,\sigma_a^2)$,
respectively, we should expect a gradient in the point color in Figures
\ref{fig:correlation}(b) and \ref{fig:correlation}(c) {\em
perpendicular} to the fitted regression.  However, we see that the
primary gradient in point color is parallel to the fitted regression.
This is consistent with the result that there is little to no
correlation between $M_i$ and $\sigma_a/v_{{\rm H}\alpha}$, as shown in
Figure \ref{fig:correlation}(d):  Both $r_s$ and $\phi$ are small and
only marginally significant with respect to their errors.  From equation
\ref{eq:dproj}, this implies that the deprojected stellar rotation is a
roughly constant fraction of the ionized-gas rotation speed, with
$v_{\ast} \sim 0.9 v_{{\rm H}\alpha}$.  Although the radius at which the
AD was sampled is different, this ratio is consistent with a similar
ratio measured for DiskMass-Survey galaxies \comment{Martinsson et al.}.

The points in Figure \ref{fig:correlation}(d) are colored by the global
$N-r$ color; however, it is difficult to determine from this
illustration if their is any correlation between $\sigma_a/v_{{\rm
H}\alpha}$ and global galaxy color.  One might expect such a correlation
if, for example, galaxies with different global color have different
light-weightings of the thin vs. thick-disk \comment{refs; not sure
there are relevant ones; Comer\'{o}n?}, which propagates to a difference
in which component dominates the dynamical pressure in the disk
midplane.  Figure \ref{fig:colortrend} assesses this directly by
plotting the error-weighted mean trend in $\sigma_a/v_{{\rm H}\alpha}$
for quartiles of the $N-r$ distribution of the AD sample.  The error
bars represent the error-weighted standard deviation of the data in each
bin, whereas the error-weighted standard error is always {\it smaller
than the plotted point}.  This suggests a marginal, yet statistically
significant, detection of a slightly larger $\sigma_a/v_{{\rm H}\alpha}$
in the reddest bin.

%; however, more data would be required to lend this result any
%statistical significance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%
\begin{center}
%
\includegraphics[width=1.0\columnwidth]{figs/adov_vs_color.pdf}
%
\end{center}
%
\caption{
%
The error-weighted mean trend in $\sigma_a/v_{{\rm H}\alpha}$ binned in
quartiles of the global $N-r$ color (black points).  The error bars
represent the error-weighted standard deviation of the data within each
bin; the error-weighted standard error is smaller than the size of each
black point.  The underlying gray histogram provides the number of
measurements in bins of $N-r$; each quartile contains $\sim$152
galaxies.
%
}
%
\label{fig:colortrend}
%
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

\subsection{Technical Concerns}

\comment{Sample bias}

\comment{Beam-smearing}

\subsection{Prospects}

\comment{If $\sigma_a$ is a direct proxy for $\sigma_R$ ...}

\bibliography{master}

\end{document}

% Under the nominal assumptions of (i) a radially independent rotation
% speed and SVE shape and (ii) a exponential dispersion profile with an
% e-folding length of twice the disk scale length ($h_R$, ref), equation
% \ref{eq:adformal} can be used to show asymmetric drift will be maximum
% at 1.25$h_R$, or approximately 0.7$R_{\rm eff}$.

%\begin{deluxetable*}{ r r r r r r r r }
%\tabletypesize{\small}
%\tablewidth{0pt}
%\tablecaption{Linear Regressions for $M_i$ [[Check slope-intercept form]]}
%\tablehead{ \colhead{Dependent} & \multicolumn{3}{c}{Parametric Form} && \multicolumn{3}{c}{Slope-intercept Form} \\
%    \cline{2-4} \cline{6-8}
%    \colhead{Variable} & \colhead{$y_0$} & \colhead{$\phi$} & \colhead{$\varepsilon$} 
%                        && \colhead{$m$} & \colhead{$b$} & \colhead{$\varepsilon_y$} }
%\startdata
% $v_{{\rm H}\alpha}$                &      2.168 &    170.70 &      0.069 &&     -0.164 &     -1.20 &      0.070 \\
%                                    & $\pm$0.003 & $\pm$0.15 & $\pm$0.001 && $\pm$0.003 & $\pm$0.05 & $\pm$0.001 \\[2pt]
% $\log(\sigma_a^2)$                 &       3.78 &     160.1 &      0.209 &&     -0.362 &     -3.69 &      0.222 \\
%                                    &  $\pm$0.01 &  $\pm$0.5 & $\pm$0.005 && $\pm$0.009 & $\pm$0.19 & $\pm$0.005 \\[2pt]
% $\log(\sigma_a/v_{{\rm H}\alpha})$ &     -0.269 &     -0.24 &      0.068 &&     -0.004 &     -0.36 &      0.068 \\
%                                    & $\pm$0.003 & $\pm$0.17 & $\pm$0.002 && $\pm$0.003 & $\pm$0.06 & $\pm$0.002
%\enddata
%\label{tab:lines}
%\end{deluxetable*}